Towards human SuperEEG

### 2018-10-03 paper/boneyard.tex
checking in more changes (reverted and revamped intro, now getting into approach)

>>>>

\section*{Introduction}
Current human brain recording techniques are fraught with compromise.
Commonly used approaches include functional magnetic resonance imaging
(fMRI), scalp electroencephalography (EEG), and magnetoencephalography
(MEG).  For each of these techniques, neuroscientists
and electrophysiologists must choose to optimize spatial resolution
at the cost of temporal resolution (e.g.\ as in fMRI) or temporal
resolution at the cost of spatial resolution (e.g.\ as in EEG
and MEG).  A less widely used approach (due to requiring work with
neurosurgical patients) is to record from electrodes implanted
directly onto the cortical surface (electrocorticography; ECoG) or
into deep brain structures (intracranial EEG; iEEG).  However, these
intracranial approaches also require compromise: the high temporal and
spatial resolutions of intracranial recordings comes at the cost of
substantially reduced brain coverage, since safety considerations
limit the number of electrodes one may implant in a given patient's
brain.  Further, the locations of implanted electrodes are determined by clinical,
rather than research, needs.

%Direct intracranial recordings are the gold standard method for
%measuring activity from deep brain structures in both humans and
%animals.
%However, one can only implant a limited number of electrodes in a
%given person's brain.  This begs the question: what can we infer
%about the activity exhibited by the rest of that person's brain,
%given what we learn from recordings taken from \textit{other people's
%brains}?
The current ``gold standard'' for precisely localizing signals and
sampling at high temporal resolution is to take (ECoG or iEEG)
recordings from implanted electrodes (but from a limited set of
locations in any given brain).  This begs the following question: what
can we infer about the activity exhibited by the rest of a person's
brain, given what we learn from the limited intracranial recordings we
have from their brain and additional recordings taken from
\textit{other} people's brains?  Here we develop an approach, based on
Gaussian process regression~\cite{Rasm06}, that uses data from
multiple people to estimate activity at arbitrary
locations in each person's brain (i.e., independent of their electrode
placements).  We test this \textit{Super EEG} approach using a large
dataset of intracranial recordings collected as neurosurgical patients
studied and recalled random word lists~\cite{SedeEtal03, SedeEtal07a,
  SedeEtal07b, MannEtal11, MannEtal12}.  We show that the Super EEG
algorithm recovers signals well from electrodes that were held out of
the training dataset.  We also examine the factors that influence how
accurately activity may be estimated (recovered), which may have important
implications for electrode design and for electrode placement in
neurosurgical applications.

%%%%%

  The correlational structure of functional neuroimaging data may be
  used to predict which moment of movie or story an individual is
  experiencing, identify which person's brain a recording came from,
  and identify the task a person is performing.  In contrast, our
  structural (anatomical) connectome is static for the duration of
  typical neuroimaging experiments, is task invariant, and is similar
  across people.  How might these seemingly opposing features of
  functional connectomes versus static structural connectomes be
  reconciled?  We developed a static task- and person-invariant
  connectome model, and explored the extent to which the model could
  explain dynamic full-brain activity patterns.  We trained the model
  using ECoG recordings from two large datasets, and we tested its
  ability to predict held-out data.  We found that neural activity
  patterns are best explained by a connectome that is common across
  individuals, but changes across tasks.

Are our brain's networks static or dynamic?  And to what extent are the network properties of our brains stable across people and tasks?  One body of work suggests that our brain's \textit{functional} networks are dynamic~\citep[e.g., ][]{MannEtal18}, person-specific~\citep[e.g., ][]{FinnEtal15}, and task-specific~\citep[e.g., ][]{Turk13}.  In contrast, although the gross anatomical structure of our brains changes meaningfully over the course of years as our brains develop, on the timescales of typical neuroimaging experiments (i.e., hours to days) our anatomical networks are largely stable~\citep[e.g., ][]{CaseEtal00}.  Further, many aspects of brain anatomy, including white matter structure, is largely preserved across people~\citep[e.g., ][]{TalaTour88, JahaEtal13, MoriEtal08}  There are several possible means of reconciling this apparent inconsistency between dynamic person- and task-specific functional networks versus stable anatomical networks.  For example, relatively small magnitude anatomical differences across people may be reflected in reliable functional connectivity differences.  Along these lines, one recent study found that diffusion tensor imaging (DTI) structural data is similar across people, but may be used to predict person-specific resting state functional connectivity data~\citep{BeckEtal18}.  Another (potentially complementary) possibility is that our functional networks are constrained by anatomy, but nevertheless exhibit rapid task-dependent changes~\citep[e.g., ][]{SporBetz16}.

Here we take a model-based approach to studying whether high
spatiotemporal resolution activity patterns throughout the human brain
may be explained by a static connectome model that is shared across
people and tasks.  Specifically, we trained a model to take in
recordings from a subset of brain locations, and then predicted
activity patterns during the same interval, but at \textit{other}
locations that were held out from the model.  Our model, based on
Gaussian process regression, relies on three assumptions (each of
which we test).  First, we assume that functional correlations are
stable over time and across tasks.  Second, we assume that some of the
correlational structure of people's brain activity is similar across
individuals.  Third, we resolve ambiguities in the data by assuming
that neural activity from nearby sources will tend to be similar, all
else being equal.  After fitting the model to an ECoG dataset, one can
then ask (for a held-out individual's brain): given what we know about
the correlational structure of \textit{other} people's brains, and
given the recordings we made from electrodes implanted in this
person's brain, how would those recordings most likely have looked at
\textit{other} locations throughout this person's brain?  We named our
general approach \textit{SuperEEG} because the trained model provides
a means of inferring full-brain activity patterns at high
spatiotemporal resolutions, using ECoG recordings taken from a limited
set of brain locations.

We tested the SuperEEG approach using two large ECoG datasets. Dataset 1 comprises multi-hour recordings from 6876 electrodes taken across several recording sessions as 88 neurosurgical patients studied random word lists periodically throughout their day~\cite{SedeEtal03, SedeEtal07a,
  SedeEtal07b, MannEtal11, MannEtal12}.  Dataset 2 comprises recordings from XXX electrodes taken as XXX patients performed a series of two memory tasks~\citep{EzzyEtal17, HoraEtal17, KragEtal17, KuceEtal17, LinEtal17, SoloEtal18, WeidEtal18, EzzyEtal18, KuceEtal18}.  In both datasets, 
  we compared the predictions made using models trained across people to predictions made using within-person data.  We found that models that incorporate data from other patients yield far more reliable predictions about an individual patient's brain activity patterns than models trained solely on that individual's data.  This indicates that at least some functional correlations are common across people.  We also used Dataset 2 to compare predictions made using data from within task, across task, or combining within and across task data.  We found that all three approaches yielded reliable predictions, but the within task predictions were the most accurate.  This indicates that some aspects of our functional connectomes are common across tasks, whereas other aspects are task-specific.  Finally, all of the models we trained (within and across patients and tasks) assumed a static connectome and yielded above-chance accuracy.  This indicates that despite moment-by-moment fluctuations in the functional connectome, at least some aspects of our connectome appear to be stable over the course of several hours.

%%%%%
